\documentclass[14pt]{extarticle}
\linespread{1.5}
\usepackage[margin=1in]{geometry}
\renewcommand{\familydefault}{\sfdefault}
\usepackage{helvet}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bbm}
\usepackage{graphicx}
\usepackage{enumitem}
\setlength{\parindent}{0pt}

\begin{document}
\subsection*{4.1 Clean \texttt{df\_demographics}}
\begin{enumerate}[label=(\alph*)]
  \item Some fields may have been missing or scarped incorrectly resulting in
    nonsensical values or \texttt{NaN} → we made sure columns have appropriate numeric data types,
    and dropped rows with missing values, or values not in expected range.
  \item All countries had valid, numeric life expectancy values, in range $[40,100]$.
  \item All countries had valid, numeric population density and urban population values.
  \item Country names are inconsistent between datasets →
    In order to keep as many countries as possible when merging datasets down the line,
    we applied a normalization scheme to country names.
  \setcounter{enumi}{5}
  \item Overall, 7 countries were affected. For Brunei, a leading space was removed.
    The rest of the mismatches were caused by irregular capitalization
    (as Python's \texttt{str.title()} does not handle words such as "and" or "of" and abbreviations).
  \begin{center}
    \begin{tabular}{ |l|l| }
      \hline
      Old value&New value\\
      \hline
      Antigua and Barbuda&Antigua And Barbuda\\
      Bosnia and Herzegovina&Bosnia And Herzegovina\\
      \ Brunei&Brunei\\
      Côte d'Ivoire&Côte D'Ivoire\\
      DR Congo&Dr Congo\\
      State of Palestine&State Of Palestine\\
      Trinidad and Tobago&Trinidad And Tobago\\
      \hline
    \end{tabular}
  \end{center}
\end{enumerate}
\begin{itemize}
  \item Rows before cleaning: 200
  \item Rows after cleaning: 200 (0 dropped)
\end{itemize}
\subsection*{4.2 Clean \texttt{df\_gdp}}
Issues encountered:
\begin{itemize}
\item Non-numeric characters (commas, currency symbols, etc.) in the
\texttt{GDP\_per\_capita\_PPP} column → prevented direct float conversion
\item Missing values introduced when malformed strings were coerced to \texttt{NaN}
→ needed to be documented and removed
\item Potential outliers in the GDP distribution → important to flag for
downstream analysis
\item Duplicate country entries → required de-duplication logic
\item Inconsistent country names → needed standardization
\end{itemize}
Actions taken:
\begin{enumerate}
\item \textbf{Type conversion}\\
Cast \texttt{GDP\_per\_capita\_PPP} to string, stripped all non-digit/decimal
characters, then converted to numeric (no invalid rows found).
\item \textbf{Missing-value handling}\\
  Exported rows where \texttt{GDP\_per\_capita\_PPP} was \texttt{NaN} to
\texttt{output/dropped\_gdp.csv}, then dropped them from the DataFrame (no rows were dropped).
\item \textbf{Outlier detection (Tukey method)}\\
Calculated Q1, Q3, and IQR; flagged values outside [Q1 – 1.5·IQR, Q3 +
1.5·IQR] as outliers (found 6 outliers).
\item \textbf{Duplicate removal}\\
  Removed duplicate rows by \texttt{Country} (no duplicates were found).
\item \textbf{Country-name standardization}\\
  Applied the same normalization as \texttt{df\_demographics}.
\end{enumerate}
\begin{itemize}
\item Rows before cleaning: 213
\item Rows after cleaning: 213 (0 dropped)
\end{itemize}
\subsection*{4.3 Clean \texttt{df\_pop}}
Issues encountered:
\begin{itemize}
\item Non-numeric characters (spaces, commas, text) in the \texttt{Population} column
→ prevented direct numeric conversion
\item Missing values created when malformed strings were coerced to \texttt{NaN} →
needed to be documented and removed
\item Potential outliers in the population distribution (on a $\log_{10}$ scale) →
important to flag for review
\item Duplicate country entries → required de-duplication logic
\item Inconsistent country names → needed standardization
\end{itemize}
Actions taken:
\begin{enumerate}
\item \textbf{Type conversion}\\
Cast Population to string, removed all non-digit/decimal characters, then
converted to numeric (no invalid rows found).
\item \textbf{Missing-value handling}\\
  Rows Population was \texttt{NaN} were dropped (no such rows found).
\item \textbf{Outlier detection (Tukey on log scale)}\\
Computed $\log_{10}$(\texttt{Population}), then calculated Q1, Q3, and IQR; flagged
values outside [Q1 – 1.5·IQR, Q3 + 1.5·IQR] as outliers (1 outlier found).
\item \textbf{Duplicate removal}\\
  Removed duplicate rows by \texttt{Country} (no duplicates were found).
\item \textbf{Country-name standardization}\\
  Applied the same normalization as \texttt{df\_demographics}.
\end{enumerate}
\begin{itemize}
\item Rows before cleaning: 260
\item Rows after cleaning: 260 (0 dropped)
\end{itemize}
\end{document}
