\documentclass[14pt]{extarticle}
\linespread{1.5}
\usepackage[margin=1in]{geometry}
\renewcommand{\familydefault}{\sfdefault}
\usepackage{helvet}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bbm}
\usepackage{graphicx}
\usepackage{enumitem}
\setlength{\parindent}{0pt}

\begin{document}
\section*{Part 1 Preview}
\subsection*{3.1 Web-crawling}
Below are the first 10 rows of \texttt{df\_demographics} after crawling the website
and casting fields to appropriate types. The crawler extracted the countries in alphabetic order,
so sorting by \texttt{Country} did not affect the output.

{\scriptsize
\begin{center}
\begin{tabular}{ |l|c|c|c|c|c|c| }
  \hline
  Country&Life Exp. (Both)&Life Exp. (Female)&Life Exp. (Male)&Urban Pop.&Urban Pop. \%&Pop. Density\\
  \hline
  Afghanistan&66.5&68.1&64.9&11704638&26.7&67\\
  Albania&80.0&81.7&78.1&1948831&70.3&101\\
  Algeria&76.7&78.1&75.3&35291973&74.4&20\\
  Angola&65.0&67.5&62.4&26847887&68.8&31\\
  Antigua and Barbuda&77.9&80.6&74.9&26823&28.5&214\\
  Argentina&77.7&80.2&75.1&44010474&96.0&17\\
  Armenia&76.0&79.7&71.8&1882155&63.8&104\\
  Aruba&76.6&79.1&74.0&48340&44.7&601\\
  Australia&84.2&86.0&82.4&23335357&86.5&4\\
  Austria&82.3&84.6&80.0&5337973&58.6&111\\
  \hline
\end{tabular}
\end{center}}

\textbf{Shape \& Column Names}\\
{\footnotesize\texttt{df\_demographics.shape: (200, 6)\\
df\_demographics.columns: ['LifeExpectancy\_Both', 'LifeExpectancy\_Female', 'LifeExpectancy\_Male', 'UrbanPopulation\_Absolute', 'UrbanPopulation\_Percentage', 'PopulationDensity']}}

\subsection*{Demographics Data Analysis}
Below are the results of runnign \texttt{describe()} on each of the numeric fields in the crawled
demographics dataset:

{\scriptsize
\begin{center}
\begin{tabular}{ |l|c|c|c|c|c|c| }
  \hline
  Stat.&Life Exp. (Both)&Life Exp. (Female)&Life Exp. (Male)&Urban Pop.&Urban Pop. \%&Pop. Density\\
  \hline
  mean             &73.98 &  76.60  &  71.38  &  23592449.77  &    57.28  &  156.02\\
  std              & 7.00 &   7.13  &   6.98  &  83188523.42  &    25.15  &  176.96\\
  min              &54.80 &  55.10  &  53.50  &         0.00  &     0.00  &    2.00\\
  25\%             &68.90 &  71.47  &  66.57  &    766036.75  &    40.22  &   35.50\\
  50\%             &74.85 &  78.05  &  71.20  &   4745905.00  &    60.55  &   88.00\\
  75\%             &79.12 &  82.03  &  76.30  &  16095833.75  &    77.43  &  222.50\\
  max              &85.80 &  88.40  &  83.10  & 956553854.00  &    98.80  &  929.00\\
  missing (or zero)& 0.00 &   0.00  &   0.00  &   10.00       &     10.00 &    0.00\\
  \hline
\end{tabular}
\end{center}}

The pearson correlation between \texttt{LifeExpectancy\_Both} and \texttt{PopulationDensity} is 0.18,
which is not considered to be a high correlation.

\subsection*{3.2 Loading the 2021 CSVs}
\textbf{GDP per Capita}\\
First 5 rows before sorting \textit{(see \texttt{output/gdp\_before\_sort.csv})}
\begin{center}
  \begin{tabular}{ |l|c| }
    \hline
    Country &\texttt{GDP\_per\_capita\_PPP}\\
    \hline
    Afghanistan &2144.1665        \\
    Albania &16353.8090           \\
    Algeria &14496.8650           \\
    Andorra &59332.2030           \\
    Angola &7408.1265             \\
    \hline
  \end{tabular}
\end{center}
First 5 rows after sorting \textit{(see \texttt{output/gdp\_after\_sort.csv} — note data was already in A→Z order)}
\begin{center}
  \begin{tabular}{ |l|c| }
    \hline
    Country &\texttt{GDP\_per\_capita\_PPP}\\
    \hline
    Afghanistan &2144.1665        \\
    Albania &16353.8090           \\
    Algeria &14496.8650           \\
    Andorra &59332.2030           \\
    Angola &7408.1265             \\
    \hline
  \end{tabular}
\end{center}
Descriptive statistics \textit{(see \texttt{output/gdp\_describe.csv})}
\begin{center}
  \begin{tabular}{ |l|c| }
    \hline
    Stat.&\texttt{GDP\_per\_capita\_PPP}\\\hline
    count & 213.000000  \\
    mean & 25822.604541 \\
    std & 25794.941595  \\
    min & 836.665600    \\
    25\% & 6214.017000  \\
    50\% & 16353.809000 \\
    75\% & 38862.090000 \\
    max & 137947.340000 \\
    \hline
  \end{tabular}
\end{center}
\textbf{Shape \& Column Names}\\
{\footnotesize\texttt{df\_gdp.shape: (213, 2)\\
df\_gdp.columns: ['Country', 'GDP\_per\_capita\_PPP']}}

\textbf{Population}\\
First 5 rows before sorting \textit{(see \texttt{output/pop\_before\_sort.csv})}
\begin{center}
\begin{tabular}{ |l|c| }
  \hline
  Country&Population\\
  \hline
  Afghanistan&40000360\\
  Africa&1413750475\\
  Africa (UN)&1413753005\\
  Albania&2849591\\
  Algeria&44761051\\
  \hline
\end{tabular}
\end{center}
First 5 rows after sorting \textit{(see \texttt{output/pop\_after\_sort.csv} — sorted A→Z)}
\begin{center}
\begin{tabular}{ |l|c| }
  \hline
  Country&Population\\
  \hline
  Afghanistan&40000360\\
  Africa&1413750475\\
  Africa (UN)&1413753005\\
  Albania&2849591\\
  Algeria&44761051\\
  \hline
\end{tabular}
\end{center}
Descriptive statistics \textit{(see \texttt{output/pop\_describe.csv} — note scientific notation for large values)}
\begin{center}
  \begin{tabular}{ |l|c| }
    \hline
    Stat.&\texttt{Population}\\\hline
    count&2.600000e+02            \\
    mean &1.687752e+08             \\
    std  &7.254974e+08              \\
    min  &5.150000e+02              \\
    25\% &5.224038e+05             \\
    50\% &6.827910e+06    \\
    75\% &3.425334e+07             \\
    max  &7.954448e+09              \\
    \hline
\end{tabular}
\end{center}
\textbf{Shape \& Column Names}\\
{\footnotesize\texttt{df\_pop.shape: (260, 2)\\
df\_pop.columns: ['Country', 'Population']}}

\section*{Part 2 Preview}
\textit{Taken from \texttt{output/cleaning\_summary.pdf}}
\subsection*{4.1 Clean \texttt{df\_demographics}}
\begin{enumerate}[label=(\alph*)]
  \item Some fields may have been missing or scarped incorrectly resulting in
    nonsensical values or \texttt{NaN} → we made sure columns have appropriate numeric data types,
    and dropped rows with missing values, or values not in expected range.
  \item All countries had valid, numeric life expectancy values, in range $[40,100]$.
  \item All countries had valid, numeric population density and urban population values.
  \item Country names are inconsistent between datasets →
    In order to keep as many countries as possible when merging datasets down the line,
    we applied a normalization scheme to country names.
  \setcounter{enumi}{5}
  \item Overall, 7 countries were affected. For Brunei, a leading space was removed.
    The rest of the mismatches were caused by irregular capitalization
    (as Python's \texttt{str.title()} does not handle words such as "and" or "of" and abbreviations).
  \begin{center}
    \begin{tabular}{ |l|l| }
      \hline
      Old value&New value\\
      \hline
      Antigua and Barbuda&Antigua And Barbuda\\
      Bosnia and Herzegovina&Bosnia And Herzegovina\\
      \ Brunei&Brunei\\
      Côte d'Ivoire&Côte D'Ivoire\\
      DR Congo&Dr Congo\\
      State of Palestine&State Of Palestine\\
      Trinidad and Tobago&Trinidad And Tobago\\
      \hline
    \end{tabular}
  \end{center}
\end{enumerate}
\begin{itemize}
  \item Rows before cleaning: 200
  \item Rows after cleaning: 200 (0 dropped)
\end{itemize}
\subsection*{4.2 Clean \texttt{df\_gdp}}
Issues encountered:
\begin{itemize}
\item Non-numeric characters (commas, currency symbols, etc.) in the
\texttt{GDP\_per\_capita\_PPP} column → prevented direct float conversion
\item Missing values introduced when malformed strings were coerced to \texttt{NaN}
→ needed to be documented and removed
\item Potential outliers in the GDP distribution → important to flag for
downstream analysis
\item Duplicate country entries → required de-duplication logic
\item Inconsistent country names → needed standardization
\end{itemize}
Actions taken:
\begin{enumerate}
\item \textbf{Type conversion}\\
Cast \texttt{GDP\_per\_capita\_PPP} to string, stripped all non-digit/decimal
characters, then converted to numeric (no invalid rows found).
\item \textbf{Missing-value handling}\\
  Exported rows where \texttt{GDP\_per\_capita\_PPP} was \texttt{NaN} to
\texttt{output/dropped\_gdp.csv}, then dropped them from the DataFrame (no rows were dropped).
\item \textbf{Outlier detection (Tukey method)}\\
Calculated Q1, Q3, and IQR; flagged values outside [Q1 – 1.5·IQR, Q3 +
1.5·IQR] as outliers (found 6 outliers).
\item \textbf{Duplicate removal}\\
  Removed duplicate rows by \texttt{Country} (no duplicates were found).
\item \textbf{Country-name standardization}\\
  Applied the same normalization as \texttt{df\_demographics}.
\end{enumerate}
\begin{itemize}
\item Rows before cleaning: 213
\item Rows after cleaning: 213 (0 dropped)
\end{itemize}
\subsection*{4.3 Clean \texttt{df\_pop}}
Issues encountered:
\begin{itemize}
\item Non-numeric characters (spaces, commas, text) in the \texttt{Population} column
→ prevented direct numeric conversion
\item Missing values created when malformed strings were coerced to \texttt{NaN} →
needed to be documented and removed
\item Potential outliers in the population distribution (on a $\log_{10}$ scale) →
important to flag for review
\item Duplicate country entries → required de-duplication logic
\item Inconsistent country names → needed standardization
\end{itemize}
Actions taken:
\begin{enumerate}
\item \textbf{Type conversion}\\
Cast Population to string, removed all non-digit/decimal characters, then
converted to numeric (no invalid rows found).
\item \textbf{Missing-value handling}\\
  Rows Population was \texttt{NaN} were dropped (no such rows found).
\item \textbf{Outlier detection (Tukey on log scale)}\\
Computed $\log_{10}$(\texttt{Population}), then calculated Q1, Q3, and IQR; flagged
values outside [Q1 – 1.5·IQR, Q3 + 1.5·IQR] as outliers (1 outlier found).
\item \textbf{Duplicate removal}\\
  Removed duplicate rows by \texttt{Country} (no duplicates were found).
\item \textbf{Country-name standardization}\\
  Applied the same normalization as \texttt{df\_demographics}.
\end{enumerate}
\begin{itemize}
\item Rows before cleaning: 260
\item Rows after cleaning: 260 (0 dropped)
\end{itemize}

\section*{Part 3 Preview}
\subsection*{5.1 New Feature}
Below is the result of running \texttt{describe()} on the new column \texttt{TotalGDP}:
\begin{center}
\begin{tabular}{ |l|r| }
  \hline
  mean&      889352331458.71\\
  std &     3095183898278.81\\
  min &         391387777.17\\
  25\% &       36903699628.76\\
  50\% &      112857767960.96\\
  75\% &      566276868415.05\\
  max &    29108919292117.52\\
  \hline
\end{tabular}
\end{center}
\subsection*{5.2+5.3 Transforms}
We applied $\log_{10}$ transformation and z-score normalization to columns\\
\texttt{GDP\_per\_capita\_PPP} and \texttt{Population}, in order to 
compress scale and center the data, providing a more comprehensible comparison.
To the column \texttt{LifeExpectancy\_Both}, which already is approximately normally distributed,
we only applied z-score normalization. Here are the results of running \texttt{describe()} on the
normalized columns:
\begin{center}
\begin{tabular}{ | l | c | c | c | }
  \hline
  Stat. & GDP & Population & Life Exp.\\
  \hline
  mean   &  -0.00 & -0.00 &-0.00\\
  std    &   1.00 &  1.00 & 1.00\\
  min    &  -2.44 & -2.39 &-2.71\\
  25\%   &  -0.76 & -0.58 &-0.74\\
  50\%   &   0.09 &  0.07 & 0.12\\
  75\%   &   0.84 &  0.71 & 0.75\\
  max    &   1.94 &  2.71 & 1.68\\
  \hline
\end{tabular}
\end{center}
\subsection*{5.4 Data Integration}
After normalizing the \texttt{Country} column across dataframes, and setting it as index, we merged them.
We were left with 173 countries:\\
\texttt{Afghanistan, Albania, Algeria, Angola, Antigua And Barbuda, Argentina, Armenia, Aruba, Australia, Austria, ...}\\
The countries \textbf{dropped} from the crawled demographics dataset are:\\
\texttt{
Cabo Verde,
Cuba,
Curaçao,
Czech Republic (Czechia),
Côte D'Ivoire,
Dr Congo,
Eritrea,
French Guiana,
French Polynesia,
Guadeloupe,
Guam,
Martinique,
Mayotte,
Micronesia,
New Caledonia,
North Korea,
Réunion,
Sao Tome \& Principe,
South Sudan,
St. Vincent \& Grenadines,
State Of Palestine,
Taiwan,
Timor-Leste,
U.S. Virgin Islands,
Venezuela,
Western Sahara,
Yemen
}
\end{document}
